---
title: "Trading Strategy"
output:
  pdf_document: default
  html_document: default
date: "2026-02-19"
header-includes:
  - \usepackage{bm}
editor_options: 
  markdown: 
    wrap: 72
---
# Introduction

In 1755 Homma Munehisa published his work "The Fountain of Gold - The Three
Monkey Record of Money" which explained his approach to trading rice futures on
the Dojima Rice Exchange. He stated that market sentiment is the critical
driver of price action, and explained his system of price analysis now called
Candlestick Charting, one of the basic techniques of Technical Analysis - the
discipline of inferring future price action from its history. 

According to orthodox financial theory, Technical Analysis should not
out-perform the market. Historical price information is available to the
overwhelming majority of market participants, both amateur and professional,
and so if any opportunity were to be gained from its analysis, it would be
acted upon and disappear very quickly. This effect is one of several inferences
drawn from the Efficient Market Hypothesis, the dominant theory of
information-driven price action in modern finance. To the Efficient Market
Hypothesis, technical analysis is little better than astrology [^2].

However, this theory is not universally accepted, and has been coming under
increased scrutiny since the early 1980s. Whilst it is certainly a good model
of what it means for a market to be efficient, and provides good grounds for
believing they should be, it does not follow that markets are efficient. For
one thing, perfectly efficient markets are themselves theoretically impossible
[^3]. Secondly, several studies of major indices indicate that even deep
markets can exhibit price autocorrelation, particularly in times of market
stress [^4]. Thirdly, the performance of Renaissance Technologies' Medallion
fund cannot be dismissed as just an anomaly [^5]. Indeed, it has been suggested
that the acceptance of the universal application of the Efficient Market
Hypothesis is directly correlated to its fashionability [^6]. 

Nonetheless, financial markets are fundamentally stochastic systems, which can
be described in terms of a drift and stochastic parameter such as:

$$
  X_t = \theta X_{t-1} + \epsilon_{t}
$$

Where $\epsilon$ is a drift coefficient, $X_t$ is the price at time $t$ and
$\epsilon$ is a random variable representing the effect of new information on
the price. An investor would like to know the signal, $\theta X_t$ and reject
$\epsilon$, the noise. 

## Technical Indicators
Modern technical analysis has many sub-disciplines, ranging from subjective 
pattern-recognition, to high-frequency, fully-automated mathematical models. At
the less subjective but still easily applied end of this spectrum lie a group
of models called "technical indicators". These are popular with amateur 
traders because they are easy to understand and apply. Indeed, many
of the most popular indicators were developed before high-power computing was
widely available and were favoured because they are easy to compute. In fact, 
most technical indicators are types of signal estimation filters used in signal
processing [^7]. Because they are all variations on the same basic mathematical
structure, they tend to correlate closely and fall into four key categories [^8]:

1. *Momentum indicators*, attempt to measure the speed of price movements.
2. *Trend indicators*, attempt to detect the current trend and its state.
3. *Volatility indicators*, indicate changes in trading activity and
   uncertainty.
4. *Volume indicators*, measure market participation. e.g. A change in price
   associated with high volume indicates a widespread conviction in the new 
   regime.

Consequently, they are usually used in combination with each other and other
techniques. In fact, indicators are often composed of other indicators. For
example, the Exponential Moving Average:

$$
  EMA_{t} = \alpha P_{t} + (1-\alpha)EMA_{t-1}
$$

is the basis of the DEMA, ZLEMA, TRIX, and ATR indicators. 

Given the above, our first query is whether or not there exists an optimal
combination of indicators that can be effectively applied without subjective
intervention, i.e. an automated trading system based on existing indicators. 

To achieve this, we utilised R's gramEvol package to generate and test
different combinations of indicators found in the TTR package. 

ChatGPT was utilised as a research tool, code analysis, and debugging, but did
not write any code. 

[^1] see, e.g. Hillier, D. (2024). eBook: Corporate Finance 5e. McGraw Hill

[^2] see Lo, A.W., Mamaysky, H. and Wang, J. (2000). Foundations of
Technical Analysis: Computational Algorithms, Statistical Inference, and
Empirical Implementation. _The Journal of Finance_, 55(4), pp.1705–1765

[^3] Grossman, S. J. and Stiglitz, J. E. (1980) 'On the Impossibility of
Informationally Efficient Markets', _The American Economic Review_, 70(3), pp.
393–408

[^4] see, e.g. Rosini, L. and Shenai, V. (2020). Stock returns and calendar
anomalies on the London Stock Exchange in the dynamic perspective of the
Adaptive Market Hypothesis: A study of FTSE100 & FTSE250 indices over a ten
year period. _Quantitative Finance and Economics_, 4(1), pp.121–147.

[^5] TODO

[^6] Gyamfi, E.N., Kyei, K.A. and Gill, R. (2017). Market efficiency of African
stock markets: A Meta-Analysis. The Journal of Developing Areas, 51(4),
pp.69–80. doi:https://doi.org/10.1353/jda.2017.0088.

[^7] See Lo et al. (ibid) and Oppenheim, A.V. and Verghese, G.C. (2017).
_Signals, systems and inference._ Harlow Pearson Education Limited.

[^8] ‌portaladmin (2025). Axis Direct. [online] Axisdirect.in. Available at:
https://simplehai.axisdirect.in/117-good-reads-ri/8935-series-22-types-of-technical-indicators
Accessed 7 Apr. 2026.

# Data

Three datasets were chosen for analysis: Berkshire-Hathaway, Apple, and Brent-Crude.
The former two are equities, which exhibit a general upward trend, and the latter
is a commodity that fluctuates within a wide range. They were selected because they
are actively traded instruments with differing returns and volatilities, representing
different market regimes.

# Trading Strategy

A context-free grammar of indicators from the TTR package for optimisation in a
Genetic Program was created. This involved wrapping the indicators to permit
appropriate error handling and a detailed guard in the fitness function to
catch the many ways that R's delightfully idiosyncratic behaviour can poison
the genetic program tree by silently propogating inappropriate values.

The indicators were selected in order to ensure that all four of the above
categories of indicator were represented.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 2.6,
  out.width = "85%",
  fig.pos = "!htbp"
)

set.seed(1230)
### Libraries #############
library(gramEvol)
library(quantmod)
library(xts)

library(PerformanceAnalytics)
library(BatchGetSymbols)
rds_dir <- "ga_models"
dir.create(rds_dir, showWarnings = FALSE)
rerun_ga <- FALSE

load_or_run_rds <- function(path, builder, rerun = rerun_ga) {
  if (!rerun && file.exists(path)) {
    readRDS(path)
  } else {
    obj <- builder()
    saveRDS(obj, path)
    obj
  }
}
### Custom Functions ##########
### Data Import #############
data_start <- "2000-01-01"
data_end   <- "2026-01-01"
tkr <- "BRK-B"

# Download into a named list (preserves order)
FullDataXTS <- getSymbols(tkr, src = "yahoo",
                          from = data_start,
                          to   = data_end,
                          auto.assign = FALSE) |> na.omit()

halfway       <- mean(as.Date(c(data_start, data_end)))
training_end  <- paste0("/", halfway)
testing_start <- paste0(as.Date(halfway) + 1, "/")
########## Change '-' in ticker names to '.' because R has a delightful habit of
########## unexpectedly sanitising the column names so they are compatible as symbols
tkr <- make.names(tkr)
colnames(FullDataXTS) <- make.names(colnames(FullDataXTS), unique = TRUE)

make_field <- \(f) paste0(tkr, ".", f)
volume <- make_field("Volume")
open   <- make_field("Open")
high   <- make_field("High")
low    <- make_field("Low")
close  <- make_field("Close")

price_aspects <- c(open, high, low, close)
price_rules <- lapply(price_aspects, as.symbol)
# Sometimes vol is 0 in commodity indexes, this gets converted into -Inf which
# will poison the entire result. Get rid of it here
vol_bad <- !is.finite(FullDataXTS[, volume])
FullDataXTS[vol_bad, volume] <- NA
FullDataXTS[, volume] <- zoo::na.approx(FullDataXTS[, volume], na.rm = FALSE)
FullDataXTS[, volume] <- zoo::na.locf(FullDataXTS[, volume], na.rm = FALSE)
FullDataXTS[, volume] <- zoo::na.locf(FullDataXTS[, volume], fromLast = TRUE)

TrainingData <- FullDataXTS[training_end]
TestingData  <- FullDataXTS[testing_start]
```

```{r indicators}
############################# Technical Analysis Indicators ####################
# To guard against inappropriate parameters or meaningless values such as NULL
# poisoning the GP process, indicators return either a valid series or an empty
# one that can be handled by other (particularly the comp) function.
mk_guarded_indic <- \(f) {
  function(data) {
    tryCatch(f(data), error = \(e) return(data[0]))
  }
}

clv <- mk_guarded_indic(TTR::CLV)
# Guard for indicators that take a price and 1 extra param
mk_guarded_indic_1 <- \(f) {
  function(data, n) {
    tryCatch(f(data, n), error = \(e) return(data[0]))
  }
}

sma   <- mk_guarded_indic_1(TTR::SMA)
ema   <- mk_guarded_indic_1(TTR::EMA)
hma   <- mk_guarded_indic_1(TTR::HMA)
vhf   <- mk_guarded_indic_1(TTR::VHF)
cmo   <- mk_guarded_indic_1(TTR::CMO)
zlema <- mk_guarded_indic_1(TTR::ZLEMA)
```
```{r guard_2, echo=FALSE}
mk_guarded_indic_2 <- \(f) {
  function(data, a, b) {
    tryCatch(f(data, a, b), error = \(e) { return(data[0])}
    )
  }
}
```
```{r indic_2}
evwma  <- mk_guarded_indic_2(TTR::EVWMA)
vwap   <- mk_guarded_indic_2(TTR::VWAP)
cmf    <- mk_guarded_indic_2(TTR::CMF)
rsi    <- mk_guarded_indic_2(TTR::RSI)
```
```{r guard_3, echo=FALSE}
mk_guarded_indic_3 <- \(f) {
  function(data, a, b, c) {
    tryCatch(f(data, a, b, c), error = \(e) return(data[0]))
  }
}
```
```{r indic_3}
alma <- mk_guarded_indic_3(TTR::ALMA)
dema  <- mk_guarded_indic_3(TTR::DEMA)
# MACD requires its own wrapper since it returns multiple columns
macd <- function(data, a, b, c, d) {
  tryCatch({
    TTR::MACD(data, a, b, c, maType = d)[, "macd"]
  }, error = function(e) data[0])
}

```
```{r comp, echo=FALSE}
# Grouped aspects for various indicators that require them
make_high_low_close <- \(high, low, close) {
  as.call(c(
    as.name("merge"),
    lapply(c(high, low, close), as.name)
  ))
}

high_low_close <- make_high_low_close(high, low, close)

comp <- function(aspect, strategy) {
  tryCatch({
    if(class(aspect)[1] == "integer") {
      signal <- xts(rep(0L, NROW(strategy)), order.by = index(strategy))
      signal[aspect > strategy] <-  1L
      signal[aspect < strategy] <- -1L
      return(signal)
    } else {
    ab <- merge(aspect, strategy, join = "left")
    # Flag rows with NA in them. This usually refers to the lag period for 
    # indicators
    ok <- is.finite(ab[, 1]) & is.finite(ab[, 2])
    signal <- xts(rep(0L, NROW(ab)), order.by = index(ab))
    # Go long
    signal[ok & ab[,1] > ab[,2]] <-  1L
    # Go short
    signal[ok & ab[,1] < ab[,2]] <- -1L
    # The series now has 1 or -1 for the signal, and 0 if the row is invalid
    signal
    }
  }, error = \(e) {
    return(aspect[0])
  }
  )
}
```

```{r indicator_rules}

make_indicator_rules <- \(price_rules, volume, high_low_close) {
  list(expr        = grule(compare(aspect, signal),
                           compare(const, signal ), 
                           compare(signal, signal ) 
                           ),
       compare     = grule(comp),
       signal      = grule(op(signal, indic),
                           op(signal, aspect),
                           indic
                           ),
       op          = grule('+', '-', '*', '/'),
       indic       = grule(sma(aspect, lag), # Trend
                           ema(aspect, lag), # Trend
                           hma(aspect, lag), # Trend
                           vhf(aspect, lag), # Trend
                           dema(aspect, lag, ratio, bool), # Trend
                           zlema(aspect, lag), # Trend
                           evwma(aspect, vol, lag), # Trend
                           vwap(aspect, vol, lag), #Volume
                           cmf(hlc, vol, lag), #Volume
                           cmo(aspect, lag), #Momentum
                           rsi(aspect, lag, maType),#Momentum
                           macd(aspect, lag, lag, lag, maType),
                           clv(hlc) # Hard to categorise; "Pressure"
                           ),
       aspect      = do.call(grule, price_rules),
       ratio       = gvrule(seq(0.1, 1.0, by = 0.1)),
       # These are commonly used window sizes in technical analysis
       lag         = gvrule(c(2L, 3L, 5L, 10L, 14L, 20L, 50L, 100L, 200L)),
       const       = gvrule(1:50),
       bool        = grule(TRUE, FALSE),
       vol         = do.call(grule, list(as.symbol(volume))),
       hlc         = do.call(grule, list(high_low_close)),
       # Can't use WMA here
       maType      = grule("EMA", "SMA", "DEMA", "ZLEMA", "HMA")
       )
}

indicator_rules <- make_indicator_rules(price_rules, volume, high_low_close)
```

The grammar is designed to create functions that produce a signal, given
context by comparison with either an aspect of the price series (the Open, High,
Low, or Close), a constant, or another signal. The signal is then converted into
a time series of 1 or -1 if it is above or below the corresponding value in the
left hand of the comparison.

```{r echo=FALSE}
forecasting_grammar <- CreateGrammar(indicator_rules)
```

The cost function assumes that the account will hold a long position whilst the
signal is 1, a short position if the signal is -1, and none if it is 0. It seeks to
select strategies that outperform buy-and-hold. However, if unconstrained it has
a tendency to select very low frequency trading strategies that randomly select
advantageous entry points in the training set, but fail to trade in the test
set. Consequently, a threshold number of trades is set as a daily rate. In order
to make the simulation more realistic and to penalise over-trading, the cost
function adds a charge for each trade of 0.1%. This is to reflect the fact that
costs are very likely to be incurred either as a topline charge or through the
spread, but are still usually very small. The final addition is a cost for
inaccuracy.

```{r cost_function}
# Create an environment with all the variables up to this point that can be 
# accessed using column names as symbols
env <- list2env(as.list(TrainingData), parent = .GlobalEnv)
target_trade_rate <- 1/21

assess_strat <- \(result, data) {
  stopifnot(is.xts(result) && is.xts(data))
  stopifnot(NROW(result) == NROW(data))
  next_day        <- stats::lag(data, -1)  |> na.omit()
  idx             <- index(result)
  next_day        <- next_day[idx] # Force the training data to align with result
  actual_returns  <- diff(log(next_day))  |> na.omit()
  actual_trend    <- sign(actual_returns)
  forecast        <- sign(result) # Just in case we need to preprocess result in future versions
  misses          <- sum(forecast != actual_trend)
  n_trades        <- sum(diff(forecast) != 0, na.rm = TRUE)
  # Remove low-trading strategies
  if (n_trades <= (NROW(data) * target_trade_rate)) return(Inf)
  buy_and_hold    <- sum(actual_returns)
  forecast        <- forecast[index(actual_returns)]
  return_path     <- as.numeric(forecast) * as.numeric(actual_returns)
  strat_return    <- sum(return_path, na.rm = TRUE)
  trade_rate      <- n_trades / length(result)
  trade_penalty   <- ((trade_rate - target_trade_rate) / target_trade_rate)^2
  fee             <- 0.01
  cost            <- (buy_and_hold - strat_return) + n_trades * fee + trade_penalty + misses/NROW(actual_trend)
  if (!is.finite(cost) || is.na(cost)) {
    Inf
  } else {
  cost
  }
}
```
 
```{r echo=FALSE}
# Lag target by 1. Since the signal will be generated on close, it cannot be
# used until the next day.
training_close  <- TrainingData[, close]
# Written by ChatGPT - the list of corner cases was beyond my experience
validate_result <- function(result, training) {
  if (length(result) == 0L) return(FALSE) # Complement of guard in EMA
  if (is.null(result)) return(FALSE)
  if (!xts::is.xts(result)) return(FALSE)
  if (length(result) == 0L || NROW(result) != NROW(training) 
      || NCOL(result) != 1) return(FALSE)
  if (all(is.na(result))) return(FALSE)
  if (any(is.infinite(as.numeric(result)))) return(FALSE)
  s <- sign(as.numeric(result))
  if (!any(s != 0, na.rm = TRUE)) return(FALSE)
  TRUE
}

indicator_fit <- \(expr) {
  result <- eval(expr, envir = env)
  if (!validate_result(result, training_close))
  {
    return(Inf)
  }
  else {
    cost <- assess_strat(result, training_close)
    if(length(cost) > 1) browser()
    cost
  }
}

monitor <- data.frame(
  iteration     = integer(),
  best_cost     = numeric(),
  finite_mean   = numeric(),
  unique_costs  = integer()
)

monitor_func <- function(result) {
  iteration     <- result$population$currentIteration
  best_cost     <- result$best$cost
  evals         <- result$population$evaluations
  finite_evals  <- evals[is.finite(evals)]
  finite_mean   <- if (length(finite_evals) > 0) {
    mean(finite_evals)
  }
  uniques       <- length(unique(evals))
  monitor       <<- rbind(
                    monitor,
                    data.frame(
                               iteration = iteration,
                               best_cost = best_cost,
                               finite_mean = finite_mean,
                               unique_costs = uniques)
  )
}
```

If `assess_strat` returns a value greater than the termination cost, 
the GP halts at that point. But since `assess_strat` reflects a strategy's 
relative out performance over buy and hold, we do not wish to cap this value. 
Consequently, the termination cost is unlimited.

```{r evolution_function, cache=TRUE}
max_iters <- 1000
main_ga_path <- file.path(rds_dir, "brk_b_main_ga.rds")
main_ga <- load_or_run_rds(main_ga_path, function() {
  monitor <<- monitor[0, ]
  ge <- GrammaticalEvolution(forecasting_grammar,
                             indicator_fit,
                             terminationCost = -Inf,
                             monitorFunc = monitor_func,
                             iterations = max_iters,
                             max.depth = 10)
  list(ge = ge, monitor = monitor)
})
ge <- main_ga$ge
monitor <- main_ga$monitor
plot(
  monitor$iteration,
  monitor$best_cost,
  type = "l",
  xlab = "Iteration",
  ylab = "Cost",
  main = "Best cost by iteration"
)

lines(
  monitor$iteration,
  monitor$finite_mean,
  col = "red"
)

legend(
  "topright",
  legend = c("Best cost", "Mean finite cost"),
  col = c("black", "red"),
  lty = 1
)

```

The above figure shows that the evolution progressed in steps, as a new
elite function was discovered and remained in top position for extended periods
whilst new generations failed to exceed its performance.

One might suspect that the GA is static during these periods, but the red line
shows that the mean cost per generation fluctuates significantly, indicating
that the GA is at work.

# Results

The strategy's performance was assessed by comparison to a buy and hold
strategy and a permutation test to assess how well the strategy captured
structurally important features of the asset's behaviour.

The permutation test creates 1000 new return paths by shuffling the daily
returns of the asset. The result is a set of random return paths within the
scope of the original asset. If the strategy's performance on the original asset
is at the extreme of the random paths, then it can be considered to be
a significant result. 

## Training Data

```{r train_evaluation}
shuffle <- \(data) { 
  xts::xts(
           sample(as.numeric(data)),
           order.by = index(data)
  )
}

e_to_s <- \(ex) paste(deparse(ex), collapse = " ")

forecasting_model  <- ge$best$expressions[1]
train_signal  <- with(TrainingData, eval(forecasting_model))
train_close   <- TrainingData[, close]
next_day      <- stats::lag(train_close, -1) |> log() |> diff() |> na.omit()
train_return  <- cumsum(train_signal * next_day)

max_test <- 1000

shuffle_train <- replicate(max_test, sum(train_signal * shuffle(next_day)))
shuffle_mean <- mean(shuffle_train)
shuffle_sd   <- sd(shuffle_train)
prob_ret     <- pnorm(shuffle_mean, shuffle_sd, lower.tail = FALSE)
```
The GA optimised to this model: `r e_to_s(forecasting_model)`


```{r plot_train_return, echo=FALSE}
train_plot <- merge(train_return, cumsum(next_day), join = "inner")
plot(index(train_plot), as.numeric(train_plot[, 1]),
     type = "l", col = "black",
     ylim = range(as.numeric(train_plot), na.rm = TRUE),
     xlab = "Date", ylab = "Cumulative return",
     main = "Training return vs buy-and-hold")
lines(index(train_plot), as.numeric(train_plot[, 2]), col = 2)
legend("topleft", legend = c("Strategy", "Buy-and-hold"),
       col = c("black", 2), lty = 1, bty = "n")
```

```{r plot_train_shuffle, echo=FALSE}
hist(shuffle_train, prob = TRUE)
lines(density(shuffle_train))
abline(v = prob_ret, col = 2)
```

### Test Data
Applying the same technique to the test period results gave the following:

```{r test_evaluation, echo=FALSE}

# Test Evaluation
test_signal        <- with(TestingData, eval(forecasting_model))

test_close    <- TestingData[, close]
test_set      <- stats::lag(test_close, -1) |> log() |> diff() |> na.omit()
test_ret      <- (test_signal * test_set) 

shuffle_test  <- replicate(max_test, sum(train_signal * shuffle(next_day)))

shuffle_mean  <- mean(shuffle_test)
shuffle_sd    <- sd(shuffle_test)
actual_ret    <- sum(test_ret)

p_value <- pnorm(actual_ret, shuffle_mean, shuffle_sd, lower.tail = FALSE)
```

```{r plot_test_return, echo=FALSE}
test_plot <- merge(cumsum(test_ret), cumsum(test_set), join = "inner")
plot(index(test_plot), as.numeric(test_plot[, 1]),
     type = "l", col = "black",
     ylim = range(as.numeric(test_plot), na.rm = TRUE),
     xlab = "Date", ylab = "Cumulative return",
     main = "BR-K Test return vs buy-and-hold")
lines(index(test_plot), as.numeric(test_plot[, 2]), col = 2)
legend("topleft", legend = c("Strategy", "Buy-and-hold"),
       col = c("black", 2), lty = 1, bty = "n")
```

```{r plot_test_shuffle, echo=FALSE}
hist(shuffle_test, prob = TRUE)
lines(density(shuffle_test))
abline(v = actual_ret, col = 2)
```
The approach was also tested against other assets, showing differing performance
in each. A sample of those tested are as follows:

### Apple
```{r AAPL, echo=FALSE}

tkr <- "AAPL"
# Download into a named list (preserves order)
FullDataXTS <- getSymbols(tkr, src = "yahoo",
                          from = data_start,
                          to   = data_end,
                          auto.assign = FALSE) |> na.omit()

halfway       <- mean(as.Date(c(data_start, data_end)))
training_end  <- paste0("/", halfway)
testing_start <- paste0(as.Date(halfway) + 1, "/")

########## Change '-' in ticker names to '.' because R has a delightful habit of
########## unexpectedly sanitising the column names so they are compatible as symbols
tkr <- make.names(tkr)
colnames(FullDataXTS) <- make.names(colnames(FullDataXTS), unique = TRUE)

volume <- make_field("Volume")
open   <- make_field("Open")
high   <- make_field("High")
low    <- make_field("Low")
close  <- make_field("Close")

price_aspects <- c(open, high, low, close)
price_rules <- lapply(price_aspects, as.symbol)
high_low_close <- make_high_low_close(high, low, close)

# Sometimes vol is 0 in commodity indexes, this gets converted into -Inf which
# will poison the entire result. Get rid of it here
vol_bad <- !is.finite(FullDataXTS[, volume])
FullDataXTS[vol_bad, volume] <- NA
FullDataXTS[, volume] <- zoo::na.approx(FullDataXTS[, volume], na.rm = FALSE)
FullDataXTS[, volume] <- zoo::na.locf(FullDataXTS[, volume], na.rm = FALSE)
FullDataXTS[, volume] <- zoo::na.locf(FullDataXTS[, volume], fromLast = TRUE)

TrainingData <- FullDataXTS[training_end]
TestingData  <- FullDataXTS[testing_start]

indicator_rules <- make_indicator_rules(price_rules, volume, high_low_close)
forecasting_grammar <- CreateGrammar(indicator_rules)
training_close <- TrainingData[, close]

# Create an environment with all the variables up to this point that can be 
# accessed using column names as symbols
env <- list2env(as.list(TrainingData), parent = .GlobalEnv)
```

Model: `r e_to_s(forecasting_model)`

```{r apple_train, echo=FALSE, cache=TRUE}
apple_ga_path <- file.path(rds_dir, "aapl_ga.rds")
ge <- load_or_run_rds(apple_ga_path, function() {
  monitor <<- monitor[0, ]
  GrammaticalEvolution(forecasting_grammar,
                       indicator_fit,
                       terminationCost = -Inf,
                       monitorFunc = monitor_func,
                       iterations = max_iters,
                       max.depth = 10)
})
```
```{r apple_stats, echo=FALSE}
forecasting_model  <- ge$best$expressions[1]
train_signal       <- with(TrainingData, eval(forecasting_model))

train_close   <- TrainingData[, close]
next_day      <- stats::lag(train_close, -1) |> log() |> diff() |> na.omit()
train_return  <- cumsum(train_signal * next_day)

shuffle_train <- replicate(max_test, sum(train_signal * shuffle(next_day)))
shuffle_mean <- mean(shuffle_train)
shuffle_sd   <- sd(shuffle_train)
prob_ret     <- pnorm(shuffle_mean, shuffle_sd, lower.tail = FALSE)

# Test Evaluation
test_signal        <- with(TestingData, eval(forecasting_model))

test_close    <- TestingData[, close]
test_set      <- stats::lag(test_close, -1) |> log() |> diff() |> na.omit()
test_ret      <- (test_signal * test_set)

shuffle_test  <- replicate(max_test, sum(train_signal * shuffle(next_day)))

shuffle_mean  <- mean(shuffle_test)
shuffle_sd    <- sd(shuffle_test)
actual_ret    <- sum(test_ret)
```
```{r aapl_plot_test_return, echo=FALSE}
test_plot <- merge(cumsum(test_ret), cumsum(test_set), join = "inner")
plot(index(test_plot), as.numeric(test_plot[, 1]),
     type = "l", col = "black",
     ylim = range(as.numeric(test_plot), na.rm = TRUE),
     xlab = "Date", ylab = "Cumulative return",
     main = "Apple Test vs buy-and-hold")
lines(index(test_plot), as.numeric(test_plot[, 2]), col = 2)
legend("topleft", legend = c("Strategy", "Buy-and-hold"),
       col = c("black", 2), lty = 1, bty = "n")

hist(shuffle_test, prob = TRUE)
lines(density(shuffle_test))
abline(v = actual_ret, col = 2)
```

### Brent Crude

```{r Brent, echo=FALSE, warning=FALSE}

tkr <- "BZ=F"

# Download into a named list (preserves order)
FullDataXTS <- getSymbols(tkr, src = "yahoo",
                          from = data_start,
                          to   = data_end,
                          auto.assign = FALSE) |> na.omit()

halfway       <- mean(as.Date(c(data_start, data_end)))
training_end  <- paste0("/", halfway)
testing_start <- paste0(as.Date(halfway) + 1, "/")

########## Change '-' in ticker names to '.' because R has a delightful habit of
########## unexpectedly sanitising the column names so they are compatible as symbols
tkr <- make.names(tkr)
colnames(FullDataXTS) <- make.names(colnames(FullDataXTS), unique = TRUE)

volume <- make_field("Volume")
open   <- make_field("Open")
high   <- make_field("High")
low    <- make_field("Low")
close  <- make_field("Close")

price_aspects <- c(open, high, low, close)
price_rules <- lapply(price_aspects, as.symbol)
high_low_close <- make_high_low_close(high, low, close)

# Sometimes vol is 0 in commodity indexes, this gets converted into -Inf which
# will poison the entire result. Get rid of it here
vol_bad <- !is.finite(FullDataXTS[, volume])
FullDataXTS[vol_bad, volume] <- NA
FullDataXTS[, volume] <- zoo::na.approx(FullDataXTS[, volume], na.rm = FALSE)
FullDataXTS[, volume] <- zoo::na.locf(FullDataXTS[, volume], na.rm = FALSE)
FullDataXTS[, volume] <- zoo::na.locf(FullDataXTS[, volume], fromLast = TRUE)

TrainingData <- FullDataXTS[training_end]
TestingData  <- FullDataXTS[testing_start]

indicator_rules <- make_indicator_rules(price_rules, volume, high_low_close)
forecasting_grammar <- CreateGrammar(indicator_rules)
training_close <- TrainingData[, close]
# Create an environment with all the variables up to this point that can be 
# accessed using column names as symbols
env <- list2env(as.list(TrainingData), parent = .GlobalEnv)
```
```{r brent_train, echo=FALSE, cache=TRUE}
brent_ga_path <- file.path(rds_dir, "brent_ga.rds")
ge <- load_or_run_rds(brent_ga_path, function() {
  monitor <<- monitor[0, ]
  GrammaticalEvolution(forecasting_grammar,
                       indicator_fit,
                       terminationCost = -Inf,
                       monitorFunc = monitor_func,
                       iterations = max_iters,
                       max.depth = 10)
})
```
```{r brent_stats, echo=FALSE}
forecasting_model  <- ge$best$expressions[1]
train_signal       <- with(TrainingData, eval(forecasting_model))

train_close   <- TrainingData[, close]
next_day      <- stats::lag(train_close, -1) |> log() |> diff() |> na.omit()
train_return  <- cumsum(train_signal * next_day)

shuffle_train <- replicate(max_test, sum(train_signal * shuffle(next_day)))
shuffle_mean <- mean(shuffle_train)
shuffle_sd   <- sd(shuffle_train)
prob_ret     <- pnorm(shuffle_mean, shuffle_sd, lower.tail = FALSE)

# Test Evaluation
test_signal        <- with(TestingData, eval(forecasting_model))

test_close    <- TestingData[, close]
test_set      <- stats::lag(test_close, -1) |> log() |> diff() |> na.omit()
test_ret      <- (test_signal * test_set)

shuffle_test  <- replicate(max_test, sum(train_signal * shuffle(next_day)))

shuffle_mean  <- mean(shuffle_test)
shuffle_sd    <- sd(shuffle_test)
actual_ret    <- sum(test_ret)
```

Model: `r e_to_s(forecasting_model)`

```{r brent_plot_test_return, echo=FALSE}
test_plot <- merge(cumsum(test_ret), cumsum(test_set), join = "inner")
plot(index(test_plot), as.numeric(test_plot[, 1]),
     type = "l", col = "black",
     ylim = range(as.numeric(test_plot), na.rm = TRUE),
     xlab = "Date", ylab = "Cumulative return",
     main = "Brent-Crude Test return vs buy-and-hold")
lines(index(test_plot), as.numeric(test_plot[, 2]), col = 2)
legend("topleft", legend = c("Strategy", "Buy-and-hold"),
       col = c("black", 2), lty = 1, bty = "n")
```
```{r brent_plot_test_shuffle, echo=FALSE}
hist(shuffle_test, prob = TRUE)
lines(density(shuffle_test))
abline(v = actual_ret, col = 2)

```
## Conclusion
The results are mixed. For some assets, our approach seems to work well, but in
others the results are less impressive. Whilst developing this report, it was
noted that there was a great deal of variation between different runs of the GA.
On some occasions a very improbable, highly profitable strategy emerged. In some
others, it produced a loser. The implication is that the GA has not fully
explored the problem space, which in turn suggests that a significantly higher
number of iterations may be required, or a deeper grammar, or a higher
`max.depth` setting, however this was not feasible with the current hardware and
time constraints.

Nonetheless, luck alone doesn't explain the performance of those strategies that
did work well. For example, the following highly successful, low probability 
strategy was found after 20,000 iterations (left overnight) on BZ=F:

```
dema(BZ.F.Open, 14L, 0.1, FALSE) + BZ.F.High - hma(BZ.F.Close, 3L)] > 
zlema(BZ.F.Close,  100L)
```

Its behaviour is entirely deterministic; why should its performance be random?
In summary, this report indicates that technical indicators may have more
predictive power than the Efficient Market Hypothesis would allow, but such an
assertion requires more research on a larger universe of assets, perhaps

# Self-Adjusting Algorithm

The following code was developed to simulate a trading system as part of the cost 
function, but was too slow to use in production. However, it could be extended to 
incrementally update the GA as it meets new data. 

```{r eval=FALSE}
strategy <- initial_strategy

closed_trades <- list()
# Any strategy will require more than 1 data point, so taking the first one at
# this point isn't going to change anything but make accumulating the historical
# data with xts.rbind a lot easier
past_data     <- data_set[1]
# Construct a trade object. The close_price should never be empty on a 
# closed trade; close_price == NULL represents a program failure
mk_trade <- \(direction, open_price, fee) {
  list(direction = direction,
       open = open_price,
       close_price = NULL,
       fee = fee
  )
}

open_position <- mk_trade(direction = 0, open_price = 0, fee = 0)

for (i in 2:NROW(data_set)) {
  ###########################UPDATE GA MODEL###################################
  if ( modulus(i, six_months) == 0) {
    new_data  <- past_data[six_months_ago:NROW(past_data)]
    ge2 <- GrammaticalEvolution(forecasting_grammar, 
                                indicator_fit, 
                                suggestions = ge$best$genome # etc...)
    strategy <- ge2$best$expresssions[1]
  }
  ############################################################################
  # Any real-world strategy must generate a signal using yesterday's data
  model <- tryCatch(
          with(past_data, eval(strategy)),
           error = function(e) {
             if (grepl("outside valid range", e$message)) {
               NULL
             } else { # have a fit
               stop(e)
             }
           }
  )
  # Signals on early data are likely to produce NA since the strategy will probably
  # use moving averages. Skip these, but keep adding to past_data
  latest_data <- data_set[i]
  # Add today's ohlcv to the past_data series
  past_data <-rbind.xts(past_data,latest_data)
  # Get the last prices as a number
  latest_close <- as.numeric(latest_data[, close])
  latest_open  <- as.numeric(latest_data[, open])
  # Skip nulls from the tryCatch
  if (is.null(model)) { next } 
  # Latest signal is the last signal in the model
  signal <- as.numeric(last(model))
  # Do some trading
  current_direction <- open_position$direction
  latest_close <- as.numeric(latest_data[, close])
  if(current_direction != signal){
    #Close the current position
    open_position$close_price <- latest_close
    closed_trades <- c(closed_trades, list(open_position))
    # Open a new position, assuming a trading fee of 2% per trade. 
    assumed_fee <- 0.02
    open_position <- mk_trade(direction   = signal,
                              open_price  = latest_close,
                              fee         = assumed_fee
    )
  } 
  # If we've reached the end of the dataset, just close the final trade
  if(i == NROW(data_set)) {
    open_position$close_price <- latest_close
    closed_trades <- c(closed_trades, list(open_position))
  }
}
trade_profit <- \(trade) (trade$direction * (trade$close_price - trade$open) - trade$fee)
total_profit <- lapply(closed_trades, trade_profit) |> as.numeric() |> sum()
print(total_profit)
```
