## Combinations of Indicators

Our next avenue of investigation was whether or not different combinations of
popular indicators could produce a reliable strategy. The strategy was to build
a context-free grammar of indicators from the TTR package for optimisation in a
Genetic Program. This involved wrapping the indicators to permit appropriate
error handling and a detailed guard in the fitness function to catch the many
ways that R's delightfully idiosyncratic behaviour can poison the genetic
program tree by silently propogating inappropriate values.  

The different indicators were selected in order to ensure all four of the 
above categories of indicator were represented. 

```{r setup echo=FALSE}
### Libraries #############
library(gramEvol)
library(quantmod)
library(xts)

library(PerformanceAnalytics)
library(BatchGetSymbols)
### Custom Functions ##########
### Data Import #############
data_start <- "2000-01-01"
data_end   <- "2026-01-01"

tkr <- "BZ=F"

# Download into a named list (preserves order)
FullDataXTS <- getSymbols(tkr, src = "yahoo",
                          from = data_start,
                          to   = data_end,
                          auto.assign = FALSE) |> na.omit()

FullDataXTS   <- log(FullDataXTS)

halfway       <- mean(as.Date(c(data_start, data_end)))
training_end  <- paste0("/", halfway)
testing_start <- paste0(as.Date(halfway) + 1, "/")

########## Change '-' in ticker names to '.' because R has a delightful habit of
########## unexpectedly sanitising the column names so they are compatible as symbols
tkr <- make.names(tkr)
colnames(FullDataXTS) <- make.names(colnames(FullDataXTS), unique = TRUE)

TrainingData <- FullDataXTS[training_end]
TestingData  <- FullDataXTS[testing_start]

### Forecasting Engine #################
# input = Price ts
# output = Forecasts, up or down, probability confidence

make_field <- \(f) paste0(tkr, ".", f)
#TODO this should use make_field
volume <- paste0(tkr, ".Volume")
open   <- paste0(tkr, ".Open")
high   <- paste0(tkr, ".High")
low    <- paste0(tkr, ".Low")
close  <- paste0(tkr, ".Close")

price_aspects <- c(open, high, low, close)
price_rules <- lapply(price_aspects, as.symbol)

# These functions wrap functions that can return NA and rturn 0 instead. Using
# Inf doesn't work since it makes the trig functions return NA.
p_log <- \(num) {
  if (any(num < 0, na.rm = TRUE)) {
    Inf
  } else {
    log(num)
  }
}

p_sin <- \(num) {
  if (any(!is.finite(num))) {
    Inf
  } else {
    sin(num)
  }
}

p_cos <- \(num) {
  if (any(!is.finite(num))) {
    Inf
  } else {
    cos(num)
  }
}

############################# Technical Analysis Indicators ####################

# To guard against inappropriate parameters or meaningless values such as NULL
# poisoning the GP process, indicators return either a valid series or an empty
# one that can be handled by other (particularly the comp) function.

# Guard for indicators that take a price and 1 extra param
mk_guarded_indic_1 <- \(f) {
  function(data, n) {
    tryCatch(f(data, n), error = \(e) return(data[0]))
  }
}

sma   <- mk_guarded_indic_1(TTR::SMA)
ema   <- mk_guarded_indic_1(TTR::EMA)
hma   <- mk_guarded_indic_1(TTR::HMA)
obv   <- mk_guarded_indic_1(TTR::OBV)
vhf   <- mk_guarded_indic_1(TTR::VHF)

mk_guarded_indic_2 <- \(f) {
  function(data, a, b) {
    tryCatch(f(data, a, b), error = \(e) { return(data[0])}
    )
  }
}

# MFI ignored since it requires a merged HLC series
# SNR ignored for same reason
# and:
# EMV, 
# SAR ignored because it requires a merged HL series
evwma  <- mk_guarded_indic_2(TTR::EVWMA)
vwap   <- mk_guarded_indic_2(TTR::VWAP)


# Although ZLEMA has a third ratio parameter, it is overridden by its second, 
# the lag which is what will be focussed on
zlema <- mk_guarded_indic_1(TTR::ZLEMA)

mk_guarded_indic_3 <- \(f) {
  function(data, a, b, c) {
    tryCatch(f(data, a, b, c), error = \(e) return(data[0]))
  }
}

alma <- mk_guarded_indic_3(TTR::ALMA)
dema  <- mk_guarded_indic_3(TTR::DEMA)

# MACD requires its own wrapper since it returns multiple columns
macd <- function(data, a, b, c, d) {
  tryCatch({
    TTR::MACD(data, a, b, c, maType = d)[, "macd"]
  }, error = function(e) data[0])
}

g_tail <- \(data, n) {
  if (n >= NROW(data)) {
    data[0]
  }
  tail(data, n)
}

comp <- function(aspect, strategy) {
  tryCatch({
    if(class(aspect)[1] == "integer") {
      signal <- xts(rep(0L, NROW(strategy)), order.by = index(strategy))
      signal[aspect > strategy] <-  1L
      signal[aspect < strategy] <- -1L
      return(signal)
    } else {
    ab <- merge(aspect, strategy, join = "left")
    # Flag rows with NA in them. This usually refers to the lag period for 
    # indicators
    ok <- is.finite(ab[, 1]) & is.finite(ab[, 2])
    signal <- xts(rep(0L, NROW(ab)), order.by = index(ab))
    # Go long
    signal[ok & ab[,1] > ab[,2]] <-  1L
    # Go short
    signal[ok & ab[,1] < ab[,2]] <- -1L
    # The series now has 1 or -1 for the signal, and 0 if the row is invalid
    signal
    }
  }, error = \(e) {
    browser()
    return(aspect[0])
  }
  )
}
```



```{r indicator_rules}
indicator_rules <- list(expr        = grule(compare(aspect, signal),
                                            compare(const,  signal),
                                            compare(signal, signal)
                                            ),
                        compare     = grule(comp),
                        signal      = grule(op(transform, indic),
                                            op(transform, aspect),
                                            indic
                                            ),
                        op          = grule('+', '-', '*', '/'),
                        indic       = grule(sma(aspect, lag),
                                            ema(aspect, lag),
                                            hma(aspect, lag),
                                            vhf(aspect, lag),
                                            obv(aspect, vol),
                                            zlema(aspect, lag),
                                            evwma(aspect, vol, lag),
                                            vwap(aspect, vol, lag),
                                            dema(aspect, lag, ratio, bool),
                                            macd(aspect, lag, lag, lag, maType)
                                            ),
                        aspect      = do.call(grule, price_rules),
                        ratio       = gvrule(seq(0.1, 1.0, by = 0.1)),
                        # These are commonly used window sizes in technical analysis
                        lag         = gvrule(c(2L, 3L, 5L, 10L, 14L, 20L, 50L, 100L, 200L)),
                        const       = gvrule(1:50),
                        bool        = grule(TRUE, FALSE),
                        vol         = do.call(grule, list(as.symbol(volume))),
                        # Can't use WMA here
                        maType      = grule("EMA", "SMA", "DEMA", "ZLEMA", "HMA")
                      )
```

The grammar is designed to create functions that produce a signal. In order to
make the signal meaningful it is given context by comparing it to either an 
aspect of the price series (the High, Low, or Close), to a constant, or to another
signal. 

The comparison function converts the signal into a time series of 1 or -1
if the signal is above or below the corresponding value in the left hand of the
comparison. 

```{r echo=FALSE}
forecasting_grammar <- CreateGrammar(indicator_rules)

g_roll <- \(data, n, f) zoo::rollapply(data, n, f)

# indicator_components <- list(expr       = grule(compare(left_side, right_side)),
#                              compare    = grule(comp),
#                              left_side  = grule(aspect, series),
#                              right_side = grule(series),
#                              scalar     = grule(op(scalar, scalar), 
#                                                 reducer(series),
#                                                 const
#                                                 ),
#                              reducer    = grule(sum, mean, sd),
#                              series     = grule(aspect,
#                                                 g_roll(aspect, const, reducer),
#                                                 op(series, series),
#                                                 ema(series, const)
#                                                 ),
#                              op         = grule('+', '-', '*', '/'),
#                              aspect     = do.call(grule, price_rules),
#                              const      = gvrule(2:200)
#                              )

# forecasting_grammar <- CreateGrammar(indicator_components)

######################## Fitness Function and GP

# Create an environment with all the variables up to this point that can be 
# accessed using column names as symbols
env <- list2env(as.list(TrainingData), parent = .GlobalEnv)

# Assume
```

The cost function assumes that the signal will hold a long position whilst the
signal is 1, and a short if the signal is -1, and none if it is 0. The cost
function seeks to outperform a buy and hold strategy for the underlying asset,
expressed simply as the return. In order to make the simulation more realistic
and to penalise over-trading, the cost function adds a charge for each trade of
0.1%. This is to reflect the fact that costs are very likely to be incurred either
as a topline charge or through the spread, but are still usually very small. The 
final addition is a cost for inaccuracy. 

```{r fitness_function}
assess_fit <- \(result, data) {
  stopifnot(is.xts(result) && is.xts(data))
  stopifnot(NROW(result) == NROW(data))
  idx             <- index(result)
  next_day        <- stats::lag(data, -1)  |> na.omit()
  next_day        <- next_day[idx] # Force the training data to align with result
  actual_returns  <- diff(next_day)       |> na.omit()
  actual_trend    <- sign(actual_returns)
  forecast        <- result # Just in case we need to preprocess result in future versions
  misses          <- forecast != actual_trend
  n_trades        <- sum(diff(forecast) != 0, na.rm = TRUE)
  buy_and_hold    <- sum(actual_returns)
  # Calculating strat return requires properly aligning the datasets on their
  # date. Since next_day is lagged, forecast should be longer. We therefore 
  # want to keep forecast's index and merge onto that. 
  # Using actual_trend * next_day causes xts to automatically align the dates,
  # leading to look-ahead bias
  aligned         <- merge(forecast, actual_returns, join = "inner")
  return_path     <- coredata(aligned[,1]) * coredata(aligned[,2])
  strat_return    <- return_path |> sum()
  l <- length(result)
  fee <- 0.001
  cost <- (buy_and_hold - strat_return) + n_trades * fee
  if (is.na(cost)) {
    Inf
  } else {
  cost
  }
}
```

 
```{r echo=FALSE}
# Lag target by 1. Since the signal will be generated on close, it cannot be
# used until the next day.
training_close  <- TrainingData[, close]

# Written by ChatGPT - the list of corner cases was beyond my experience
validate_result <- function(result, training) {
  if (length(result) == 0L) return(FALSE) # Complement of guard in EMA
  if (is.null(result)) return(FALSE)
  if (!xts::is.xts(result)) return(FALSE)
  if (NROW(result) != NROW(training)) return(FALSE)
  if (NCOL(result) != 1) return(FALSE)
  if (NROW(result) == 0) return(FALSE)
  if (all(is.na(result))) return(FALSE)
  if (any(is.infinite(as.numeric(result)))) return(FALSE)
  s <- sign(as.numeric(result))
  if (!any(s != 0, na.rm = TRUE)) return(FALSE)
  TRUE
}

indicator_fit <- \(expr) {
  result <- eval(expr, envir = env)
  if (!validate_result(result, training_close))
  {
    return(Inf)
  }
  else {
    # if (sd(result, na.rm = TRUE) < 1e-6) return(Inf) # Don't allow very stable results through - they will just copy the asset
    cost <- assess_fit(result, training_close)
    if(length(cost) > 1) browser()
    cost
  }
}
```

Finally, the evolution function is called with an unlimited termination cost. 
If the `assess_fit` returns a value greater than the termination cost, then 
the GP halts at that point. But since assess_fit is dominated by a strategies 
relative outperformance over buy and hold, we do not wish to cap this value. Hence, 
it is set to `-Inf`.

```{r evolution_function}
# Run
ge <- GrammaticalEvolution(forecasting_grammar,
                           indicator_fit,
                           terminationCost = -Inf,
                           verbose = FALSE,
                           iterations = 1000,
                           max.depth = 10)
```



```{r evaluation }
# Evaluation
ForecastingModel  <- ge$best$expressions[1]
result            <- with(TestingData, eval(ForecastingModel))
error             <- assess_fit(result, TestingData[, close])

test_set      <- stats::lag(TestingData[, close], -1) |> diff() |> na.omit()
common_dates  <- intersect(index(result), index(test_set))
result        <- result[common_dates]
test_set      <- test_set[common_dates]
signal        <- (result * test_set) |> cumsum()

plot(signal)
lines(cumsum(test_set), col = 2)
print(ForecastingModel)

long_only_signal <- cumsum(test_set)
model_signal <- cumsum(sign(result) * test_set)
no_shorts_signal <- cumsum(ifelse(sign(result) > 0, 1, 0) * test_set)


################ Permutation Test 

sig <- sign(zoo::na.fill(stats::lag(result, 1), 0))

shuffle <- \(data) { 
  xts::xts(
           sample(as.numeric(data)),
           order.by = index(data)
  )
}

actual_pnl    <- sig * test_set
shuffled_data <- numeric(0)

max_test <- 1000

for (i in 1:max_test) {
  shuffled_pnl <- sum(sig * shuffle(test_set))
  shuffled_data <- c(shuffled_data, shuffled_pnl)
}

shuffle_mean  <- mean(shuffled_data)
shuffle_sd    <- sd(shuffled_data)
actual_ret    <- sum(actual_pnl)

p_value <- pnorm(actual_ret, shuffle_mean, shuffle_sd, lower.tail = FALSE)
print(p_value)

plot(cumsum(actual_pnl), col = "black")
lines(cumsum(shuffled_pnl), col = "blue")
lines(cumsum(test_set), col = "red")

cat("Model ", str(ForecastingModel))
```
