---
title: "Trading System"
output:
  pdf_document: default
  html_document: default
date: "2026-03-31"
header-includes:
  - \usepackage{bm}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(quantmod)
library(gramEvol)
```

# Introduction

In 1755 Homma Munehisa published his work "The Fountain of Gold - The Three
Monkey Record of Money" which explained his approach to trading rice futures on
the Dojima Rice Exchange. In it he stated that market sentiment is the critical
driver of price action, and explained his system of price analysis now called
"Candlestick Charting", one of the basic techniques of Technical Analysis - the
discipline of inferring future price action from its history. 

According to orthodox financial theory, Technical Analysis should not
out-perform the market. Historical price information is available to the
overwhelming majority of market participants, both amateur and professional,
and so if any opportunity were to be gained from its analysis, it would be
acted upon and disappear very quickly. This effect is one of several inferences
drawn from the Efficient Market Hypothesis, the dominant theory of
information-driven price action in modern finance[^1]. To the Efficient Market
Hypothesis's adherents, technical analysis is little better than astrology
[^2].

However, this theory is not universally accepted, and has been coming under
increased scrutiny since the early 1980s. Whilst it is certainly a good model
of what it means for a market to be efficient, and provides good grounds for
believing they should be, it does not follow that markets are efficient. For
one thing, perfectly efficient markets are themselves theoretically impossible
[^3]. Secondly, several studies of major indices indicate that even deep
markets can exhibit price autocorrelation, particularly in times of market
stress [^4]. Thirdly, the performance of Renaissance Technologies' Medallion
fund cannot be dismissed as just an anomaly [^5]. Indeed, it has been suggested
that the acceptance of the universal application of the Efficient Market
Hypothesis is directly correlated to its fashionability [^6]. 

Nonetheless, financial markets are fundamentally stochastic systems, which can
be described in terms of a drift and stochastic parameter such as:

$$
  X_t = \theta X_{t-1} + \epsilon_{t}
$$

Where $\epsilon$ is a drift coefficient, $X_t$ is the price at time $t$ and
$\epsilon$ is a random variable representing the effect of new information on
the price. The question, therefore, is whether or not $\epsilon$ is a set of 
identical, independently distributed, random variables, or not.

## Technical Indicators
Modern technical analysis has many sub-disciplines, ranging from subjective 
pattern-recognition, to high-frequency, fully-automated mathematical models. At
the less subjective but still easily applied end of this spectrum lie a group
of models called "technical indicators". These are popular with amateur 
technical analysts because they are easy to understand and apply. Indeed, many
of the most popular indicators were developed before high-power computing was
widely available and were favoured because they are easy to compute. In fact, 
most technical indicators are types of signal estimation filters used in signal
processing [^7]. Because they are all variations on the same basic mathematical
structure, and applied to the same datasets, they tend to correlate closely and
fall into four key categories [^8]:

1. *Momentum indicators*, which attempt to measure the speed of price movements.
2. *Trend indicators*, which attempt to detect the current trend and its state.
3. *Volatility indicators*, which indicate changes in trading activity and
   uncertainty.
4. *Volume indicators*, which measure market participation. e.g. A change in price
   associated with high volume indicates a widespread conviction in the new 
   regime.

Consequently, they are usually used in combination with each other and other
techniques such as candlestick analysis. In fact, indicators are often
composed of other indicators. For example, the Exponential Moving Average:

$$
  EMA_{t} = \alpha P_{t} + (1-\alpha)EMA_{t-1}
$$

is the basis of the DEMA, ZLEMA, TRIX, and ATR indicators. 

Given the above, our first query is whether or not there exists an optimal
combination of indicators that can be effectively applied without subjective
intervention, i.e. an automated trading system based on existing indicators. 

Our second query was, given their homogenous nature and shared underpinnings, 
whether or not a new indicator could be found by combining the shared mathematical
components of these existing indicators in a new way. 

To achieve this, we utilised R's gramEvol package to generate and test
different combinations of indicators found in the TTR package. We also
attempted to create a new indicator from the mathematical components of these
indicators.

ChatGPT was utilised as a research tool, code analysis, and debugging, but did
not write any code. 

[^1] see, e.g. Hillier, D. (2024). eBook: Corporate Finance 5e. McGraw Hill

[^2] see Lo, A.W., Mamaysky, H. and Wang, J. (2000). Foundations of
Technical Analysis: Computational Algorithms, Statistical Inference, and
Empirical Implementation. _The Journal of Finance_, 55(4), pp.1705–1765

[^3] Grossman, S. J. and Stiglitz, J. E. (1980) 'On the Impossibility of
Informationally Efficient Markets', _The American Economic Review_, 70(3), pp.
393–408

[^4] see, e.g. Rosini, L. and Shenai, V. (2020). Stock returns and calendar
anomalies on the London Stock Exchange in the dynamic perspective of the
Adaptive Market Hypothesis: A study of FTSE100 & FTSE250 indices over a ten
year period. _Quantitative Finance and Economics_, 4(1), pp.121–147.

[^5] TODO

[^6] Gyamfi, E.N., Kyei, K.A. and Gill, R. (2017). Market efficiency of African
stock markets: A Meta-Analysis. The Journal of Developing Areas, 51(4),
pp.69–80. doi:https://doi.org/10.1353/jda.2017.0088.

[^7] See Lo et al. (ibid) and Oppenheim, A.V. and Verghese, G.C. (2017).
_Signals, systems and inference._ Harlow Pearson Education Limited.

[^8] ‌portaladmin (2025). Axis Direct. [online] Axisdirect.in. Available at:
https://simplehai.axisdirect.in/117-good-reads-ri/8935-series-22-types-of-technical-indicators
\[Accessed 7 Apr. 2026\].

## Price Forecasting

Initially, we attempted to use gramEvol's genetic programming algorithm to 
determine the price of a stock one period ahead. This strategy assumed that the
price of a stock could be deduced from earlier prices. 

10 stocks were tested using their simple prices, log returns, and differenced
log returns with the following fitness function and grammar:

```{r forecasting}

# Grammar - note that the data rule, which references the columns of the lagged
# dataset is introduced during processing
ForecastingRules <- list(expr       = grule(op(arithmetic, arithmetic),
                                            reducer(lists),
                                            func(arithmetic),
                                            data
                                            ),
                         arithmetic = grule(op(arithmetic, arithmetic), # Things that are acceptable to an arithmetic function in R
                                            func(arithmetic), # Only apply the general functions to scalars
                                            reducer(lists), # reducers will transform a scalar into a list
                                            stats(data), 
                                            stats2(data, data),
                                            const),
                         func       = grule(p_sin, p_cos, exp, p_log),
                         op         = grule('+', '-', '*', '/'),
                         reducer    = grule(rowMeans),
                         stats      = grule(var_scalar, mean),
                         stats2     = grule(cor_scalar),
                         lists      = grule(merge(data, lists), merge(data, data)),
                         const      = gvrule(seq(0, 200, by = 0.1))
                         )

# Fitness function (RMSE)
fitness <- forecastingfitnessRMSE <- function(env, training, expr) {
  result <- eval(expr, envir = env)
  if (any(is.nan(result)))
    Inf
  else
    sqrt(mean((training - result)^2))
}
```

# TODO put the full code in here, echo off

The mean squared errors and expressions obtained were as follows:

|-------|----------------------|-----------------------|
|Ticker | Simple (RMSE)        | log (RMSE)            |
|-------|----------------------|-----------------------|
| AAPL  | High Lag 1 (2.59)    | Open Lag 3 (1.03)     | 
|-------|----------------------|-----------------------|
| BRK-B | Close Lag 1 (3.054)  | High Lag 2 (1.02)     |
|-------|----------------------|-----------------------|
| KGF.L | Close Lag 2 (7.38)   | Close Lag 5 (1.03)    |
|-------|----------------------|-----------------------|
| MKS.L | Close Lag 1 (5,43)   | Close Lag 2 (1.03)    |
|-------|----------------------|-----------------------|
| AHT.L | Close Lag 1 (66.88)  | Close Lag 1 (1.02)    |
|-------|----------------------|-----------------------|

Note that each of these results is a very recent lag. In modern financial
theory financial assets are understood to frequently behave like martingales,
particularly at higher time frames due to market efficiency. Martingales have
the particular property that $E[X_t+1 | X_{t}, X_{t-1} ... X_n] = X_{t}]$ or,
more prosaically, the best forecast of tomorrow's price is today's. The above
result strongly suggests that these stocks, and no doubt many others, are
martingales, which suggests that the price series alone cannot be used to
forecast future prices. We need an algorithm that can deduce an assets future
behaviour, rather than a precise forecast. 

# Technical Analysis


