---
title: "Trading System"
output:
  pdf_document: default
  html_document: default
date: "2026-03-31"
header-includes:
  - \usepackage{bm}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(quantmod)
library(gramEvol)
```

# Introduction

In 1755 Homma Munehisa published his work "The Fountain of Gold - The Three
Monkey Record of Money" which explained his approach to trading rice futures on
the Dojima Rice Exchange. In it he stated that market sentiment is the critical
driver of price action, and explained his system of price analysis now called
"Candlestick Charting", one of the basic techniques of Technical Analysis - the
discipline of inferring future price action from its history. 

According to orthodox financial theory, such systems should not out-perform the
market. Historical price information is available to the overwhelming majority
of market participants, both amateur and professional, and so if any
opportunity were to be gained from its analysis, it would be acted upon and
disappear very quickly. This effect is one of several inferences drawn from the
Efficient Market Hypothesis, the dominant theory of information-driven price
action in modern finance[^1]. To the Efficient Market Hypothesis's adherents,
technical analysis is little better than astrology [^2].

However, this theory is not universally accepted, and has been coming under
increased scrutiny since the early 1980s. Whilst it is certainly a good model
of what it means for a market to be efficient, it does not follow that all
markets are always efficient. For one thing, it is theoretically impossible
that it is be absolutely true at all times [^3]. Secondly, several studies of
major indices indicate that even deep markets can exhibit price
autocorrelation, particularly in times of market stress [^4]. Thirdly, the
performance of Renaissance Technologies' Medallion fund cannot be dismissed as
just an anomaly [^5]. Indeed, it has been suggested that the plausibility of the
Efficient Market Hypothesis is directly correlated to its fashionability [^6].
In particular, Lo et al (ibid) provide evidence supporting the predictive power
of several technical patterns. 

Nonetheless, financial markets are fundamentally stochastic systems, which can
be described in terms of a drift and stochastic parameter such as:

$$
  X_t = \theta X_{t-1} + \epsilon_{t}
$$

Where $\epsilon$ is a drift coefficient, $X_t$ is the price at time $t$ and
$\epsilon$ is a random variable representing the effect of new information on
the price. The question, therefore, is whether or not $\epsilon$ is a set of 
identical, independently distributed, random variables, or not.

## Modern Technical Analysis
Modern 

[^1] see, e.g. Hillier, D. (2024). eBook: Corporate Finance 5e. McGraw Hill

[^2] see Lo, A.W., Mamaysky, H. and Wang, J. (2000). Foundations of
Technical Analysis: Computational Algorithms, Statistical Inference, and
Empirical Implementation. _The Journal of Finance_, 55(4), pp.1705–1765

[^3] Grossman, S. J. and Stiglitz, J. E. (1980) 'On the Impossibility of
Informationally Efficient Markets', The American Economic Review, 70(3), pp.
393–408

[^4] see, e.g. Rosini, L. and Shenai, V. (2020). Stock returns and calendar
anomalies on the London Stock Exchange in the dynamic perspective of the
Adaptive Market Hypothesis: A study of FTSE100 & FTSE250 indices over a ten
year period. _Quantitative Finance and Economics_, 4(1), pp.121–147.

[^5] 

[^6] Gyamfi, E.N., Kyei, K.A. and Gill, R. (2017). Market efficiency of African
stock markets: A Meta-Analysis. The Journal of Developing Areas, 51(4),
pp.69–80. doi:https://doi.org/10.1353/jda.2017.0088.

## Price Forecasting

Initially, we attempted to use gramEvol's genetic programming algorithm to 
determine the price of a stock one period ahead. This strategy assumed that the
price of a stock could be deduced from earlier prices. 

10 stocks were tested using their simple prices, log returns, and differenced
log returns with the following fitness function and grammar:

```{r forecasting}

# Grammar - note that the data rule, which references the columns of the lagged
# dataset is introduced during processing
ForecastingRules <- list(expr       = grule(op(arithmetic, arithmetic),
                                            reducer(lists),
                                            func(arithmetic),
                                            data
                                            ),
                         arithmetic = grule(op(arithmetic, arithmetic), # Things that are acceptable to an arithmetic function in R
                                            func(arithmetic), # Only apply the general functions to scalars
                                            reducer(lists), # reducers will transform a scalar into a list
                                            stats(data), 
                                            stats2(data, data),
                                            const),
                         func       = grule(p_sin, p_cos, exp, p_log),
                         op         = grule('+', '-', '*', '/'),
                         reducer    = grule(rowMeans),
                         stats      = grule(var_scalar, mean),
                         stats2     = grule(cor_scalar),
                         lists      = grule(merge(data, lists), merge(data, data)),
                         const      = gvrule(seq(0, 200, by = 0.1))
                         )

# Fitness function (RMSE)
fitness <- forecastingfitnessRMSE <- function(env, training, expr) {
  result <- eval(expr, envir = env)
  if (any(is.nan(result)))
    Inf
  else
    sqrt(mean((training - result)^2))
}
```

# TODO put the full code in here, echo off

The mean squared errors and expressions obtained were as follows:

|-------|----------------------|-----------------------|
|Ticker | Simple (RMSE)        | log (RMSE)            |
|-------|----------------------|-----------------------|
| AAPL  | High Lag 1 (2.59)    | Open Lag 3 (1.03)     | 
|-------|----------------------|-----------------------|
| BRK-B | Close Lag 1 (3.054)  | High Lag 2 (1.02)     |
|-------|----------------------|-----------------------|
| KGF.L | Close Lag 2 (7.38)   | Close Lag 5 (1.03)    |
|-------|----------------------|-----------------------|
| MKS.L | Close Lag 1 (5,43)   | Close Lag 2 (1.03)    |
|-------|----------------------|-----------------------|
| AHT.L | Close Lag 1 (66.88)  | Close Lag 1 (1.02)    |
|-------|----------------------|-----------------------|

Note that each of these results is a very recent lag. In modern financial
theory financial assets are understood to frequently behave like martingales,
particularly at higher time frames due to market efficiency. Martingales have
the particular property that $E[X_t+1 | X_{t}, X_{t-1} ... X_n] = X_{t}]$ or,
more prosaically, the best forecast of tomorrow's price is today's. The above
result strongly suggests that these stocks, and no doubt many others, are
martingales, which suggests that the price series alone cannot be used to
forecast future prices. We need an algorithm that can deduce an assets future
behaviour, rather than a precise forecast. 

# Technical Analysis


