---
title: "Trading System"
output:
  pdf_document: default
  html_document: default
date: "2026-03-31"
header-includes:
  - \usepackage{bm}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(quantmod)
library(gramEvol)
```
# Introduction

## Price Forecasting
Initially, we attempted to use gramEvol's genetic programming algorithm to 
determine the price of a stock one period ahead. This strategy assumed that the
price of a stock could be deduced from earlier prices. 

10 stocks were tested using their simple prices, log returns, and differenced
log returns with the following fitness function and grammar:

```{r forecasting}

# Grammar - note that the data rule, rhich references the columns of the lagged
# dataset is introduced during processing
ForecastingRules <- list(expr       = grule(op(arithmetic, arithmetic),
                                            reducer(lists),
                                            func(arithmetic),
                                            data
                                            ),
                         arithmetic = grule(op(arithmetic, arithmetic), # Things that are acceptable to an arithmetic function in R
                                            func(arithmetic), # Only apply the general functions to scalars
                                            reducer(lists), # reducers will transform a scalar into a list
                                            stats(data), 
                                            stats2(data, data),
                                            const),
                         func       = grule(p_sin, p_cos, exp, p_log),
                         op         = grule('+', '-', '*', '/'),
                         reducer    = grule(rowMeans),
                         stats      = grule(var_scalar, mean),
                         stats2     = grule(cor_scalar),
                         lists      = grule(merge(data, lists), merge(data, data)),
                         const      = gvrule(seq(0, 200, by = 0.1))
                         )

# Fitness function (RMSE)
fitness <- forecastingfitnessRMSE <- function(env, training, expr) {
  result <- eval(expr, envir = env)
  if (any(is.nan(result)))
    Inf
  else
    sqrt(mean((training - result)^2))
}
```

# TODO put the full code in here, echo off

The mean squared errors and expressions obtained were as follows:

|-------|----------------------|-----------------------|
|Ticker | Simple (RMSE)        | log (RMSE)            |
|-------|----------------------|-----------------------|
| AAPL  | High Lag 1 (2.59)    | Open Lag 3 (1.03)     | 
|-------|----------------------|-----------------------|
| BRK-B | Close Lag 1 (3.054)  | High Lag 2 (1.02)     |
|-------|----------------------|-----------------------|
| KGF.L | Close Lag 2 (7.38)   | Close Lag 5 (1.03)    |
|-------|----------------------|-----------------------|
| MKS.L | Close Lag 1 (5,43)   | Close Lag 2 (1.03)    |
|-------|----------------------|-----------------------|
| AHT.L | Close Lag 1 (66.88)  | Close Lag 1 (1.02)    |
|-------|----------------------|-----------------------|

Note that each of these results is a very recent lag. In modern financial
theory financial assets are understood to frequently behave like martingales,
particularly at higher time frames due to market efficiency. Martingales have
the particular property that $E[X_t+1 | X_{t}, X_{t-1} ... X_n] = X_{t}]$ or,
more prosaically, the best forecast of tomorrow's price is today's. The above
result strongly suggests that these stocks, and many if not all others, are
martingales, which strongly suggests that the price series alone cannot be used
to forecast future prices. What we need is a algorithm that can deduce an
assets future behaviour, rather than a precise forecast. 

# Technical Analysis


